---
title: "Practical Machine Learning"
author: "Chetana T"
date: "16 February 2018"
output: html_document
---

## Overview
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

 To use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
 
The goal of the project is to predict the manner in which they did the exercise.

## Getting and loading the data

```{r}
setwd("c:/Users/drkeshwala/Desktop/coursera/practical Machine learning")
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"  
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

## On inspection in Excel, found NA,#DIV/0! and blank values in the data. These are not valid observed values, so remove with na.strings parameter.

training <- read.csv(url(trainUrl), na.strings = c("NA","#DIV/0!","")) 
testing <- read.csv(url(testUrl), na.strings= c("NA","#DIV/0!",""))
dim(training)
dim(testing)
```

```{r}
library(caret)
library(rpart)
library(rattle)
library(ggplot2)
set.seed(12345)
```

## Data Cleansing

Removing the non-predicators from the training set.
```{r}
training.f <- training[,-c(1:7)]
testing.f <- testing[, -c(1:7)]
dim(training.f)
dim(testing.f)
```

Removing the Na's from the dataset.
```{r}
training.f <- training.f[,colSums(is.na(training.f)) == 0]
dim(training.f)
testing.f<- testing.f[,colSums(is.na(testing.f)) == 0]
dim(testing.f)
```

## Divide the data into training and testing set.
```{r}
inTrain <- createDataPartition(training.f$classe, p = 0.6, list = FALSE)
in_training <- training.f[inTrain,]
in_testing <- training.f[-inTrain,]
dim(in_training)
dim(in_testing)
```

## Algorithm

### Using the Decision Tree algorithm for prediction.
```{r}
modFit1 <- train( classe~., method = "rpart", data = in_training)
fancyRpartPlot(modFit1$finalModel)
```

```{r}
prediction_dt <- predict(modFit1, in_testing)
confusionMatrix(prediction_dt,in_testing$classe)
```

From the Decision Tree, the Accuracy value is "0.4994".

### using the RandomfOrest algorithm for prediction.

```{r}
modFit2 <- train(classe~., data = in_training, method = "rf",
                trControl = trainControl(method = "cv",5))
prediction_rf <-predict(modFit2,in_testing)
confusionMatrix(prediction_rf,in_testing$classe)
```

using the RandomForest algorithm, the accuracy value is "0.9934".

## The important predictors in the model
```{r}
varImp(modFit2)
```

##  Conclusion

```{r}
values <- predict(modFit2,newdata = testing.f)
values
```
